Semi-quantitative petrographic characterization of Bronze Age quartzite tools – an exploratory approach

### 2023-01-09 analysis/paper/paper.Rmd
Updated relative paths for subscripts to use "here" package, removed old texts and code chunks that were greyed out

#Import descriptive metadata
metadata.csv <-
  read.csv2("./analysis/data/raw_data/metadata.csv", sep = ";", header = TRUE, na = c("", "NA", "NULL"), encoding = "UTF-8")

@@

```{r distribution-map, fig.width=15, fig.height=15, fig.fullwidth=TRUE, fig.cap="Geographical distribution of the sampled sites along the river of Ångermanälven, Västerbotten, Sweden.The glacial disturbance is visible in the widely distributd moraine. Made with ©SGU, Lantmäteriet (CC0), Natural Earth, and Riksantikvarieämbetet."}
knitr::include_graphics('../figures/site_distribution_soils_annotated.png', error = FALSE)
```

@@

```{r sample-hue-pictures, fig.cap= "An assortment of points demonstrating the variation in color and heterogeneity of the material."}
knitr::include_graphics('../figures/sample_hue.jpg', error = FALSE)
```

@@

\linebreak
<!--
## O2PLS

O2PLS is a multivariate regression method used to integrate data from different sources into a single model [@trygg02]. It builds on the Orthogonal Partial Least Squares (OPLS) method, which was developed to correct for systematic noise in a single dataset and identify latent structures [@bylesjo_etal07]. The O2PLS method takes the systematic variation in two datasets and divides it into two parts, joint variation between the two blocks of data, and the orthogonal variation that is block specific. This makes it possible to test the potential overlap between separate instrumentation, in this case NIRS and XRF.

O2PLS was implemented in R [@rcoreteam21] using the *OmicsPLS* package [@bouhaddani_etal18]. The data was pre-processed according to the above descriptions, and then cross-validation was used in order to determine the number of components to fit to the model. 
-->
\pagebreak

@@

```{r nir-spectra-summary, fig.cap= "Summary of the raw spectra for each colour group. Note the differences in absorbance between the groups, with the dark group featuring a larger range of absorbance values as well as less distinct peaks.", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/001-nir-spectra-summary.R"), echo = FALSE}
quietly(source("./analysis/paper/001-nir-spectra-summary.R", encoding = "UTF-8"))
knitr::include_graphics('../figures/001-nir-spectra-summary.png', error = FALSE)
```

@@

```{r nir-savitzky-transform, fig.cap= "Comparison of NIR spectra between two samples from the dark group. a) the non-processed spectra for sample 258 and 411, b) sample 258 spectra with 2^nd^ derivative Savitzky-Golay filtering, c) sample 411 spectra with 2^nd^ derivative Savitzky-Golay filtering.", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/003-nir-savitzky-transform.R"), echo = FALSE}
quietly(source("./analysis/paper/003-nir-savitzky-transform.R", encoding = "UTF-8"))
knitr::include_graphics("../figures/003-nir-savitzky-transform.png", error = FALSE)
```

@@

```{r nir-savitzky-transform-alt, fig.cap= "SG transformed spectra contrasting the results between the spectra in the dark (black) and light (orange) groups.", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/003-nir-savitzky-transform-alt.R"), echo = FALSE}
quietly(source("./analysis/paper/003-nir-savitzky-transform-alt.R", encoding = "UTF-8"))
knitr::include_graphics("../figures/003-nir-savitzky-transform-alt.png", error = FALSE)
```

```{r nir-peak-summary, fig.cap= "The main features in the NIR spectra can be related to overtones and vibration combinations of OH groups. The 1 400 nm and 1 900 nm peaks are indicative of molecular water, and common in quartz-based materials. The 2 200 nm and 2 350 nm bands indicate metal-OH combinations, where a more intense 2 200 nm band has typically been related to the presence of AlOH, and a more intense 2 350 nm band MgOH. Although difficult to detect in the non-processed spectra, weaker features also appear at 1 065 nm and 1 300 nm, that may be the result of ferrous iron (Fe^2+^)", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/002-nir-peak-summary.R"), echo = FALSE}
quietly(source("./analysis/paper/002-nir-peak-summary.R", encoding = "UTF-8"))
knitr::include_graphics('../figures/002-nir-peak-summary.png', error = FALSE)
```

@@

```{r nir-pca, fig.width=20, fig.height=20, fig.cap= "PCA score plot", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/004-nir-pca.R"), echo = FALSE}
quietly(source("./analysis/paper/004-nir-pca.R", encoding = "UTF-8"))
nir.pc1var <- nir.pc1var
nir.pc2var <- nir.pc2var
nir.pc3var <- nir.pc3var
```

@@

```{r nir-pca-fig, fig.width=20, fig.height=20, fig.cap= "NIR PCA score plot for PC 1-3: a) dark and light material is separated out along PC 1, due to absorbance, with white and colourless material separating along PC 2, b) PC 3 adds little new information, mainly separating out a couple of light samples.", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/004-nir-pca.R"), echo = FALSE}
#quietly(source("./analysis/paper/004-nir-pca-alt.R", encoding = "UTF-8"))
knitr::include_graphics("../figures/004-nir-pca-with-photos.png", error = FALSE)
```

@@

```{r raman-spectra-summary, fig.cap= "The non-processed Raman spectra visualised by colour group. Note that the dark material in general feature more intense peaks at wavenumber ~ 1 300 - 1 600 (cm^-1^).", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/005-raman-spectra-summary.R"), echo = FALSE}
quietly(source("./analysis/paper/005-raman-spectra-summary.R", encoding = "UTF-8"))
knitr::include_graphics("../figures/005-raman-spectra-summary.png", error = FALSE)
```

@@

```{r raman-peak-ident, fig.cap= "The two features that show up in the Raman spectra can be related to quartz and carbon structures in the material. The initial three peaks in the 100 - 500 cm^-1^ range correspond to some of the more well-known vibration modes of quartz. The 1 287 cm-^-1^ and 1 598 cm-^-1^ peaks fall within what is referred to as the D and G band respectively, indicating the presence of carbon structures.", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/006-raman-peak-identification.R"), echo = FALSE}
quietly(source("./analysis/paper/006-raman-peak-identification.R", encoding = "UTF-8"))
knitr::include_graphics("../figures/006-raman-peak-identification.png", error = FALSE)
```

@@

```{r raman-pca-var, cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/007-raman-pca.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/007-raman-pca.R", encoding = "UTF-8")))
raman.pc1var <- raman.pc1var
raman.pc2var <- raman.pc2var
raman.pc3var <- raman.pc3var
```

@@

```{r raman-pca, fig.cap= "PCA score plot for Raman data. Objects that feature only quartz peaks are clustered towards the top left of the plot, whilst the darker material that feature dominant D/G peaks are distributeted towards the right of the plot. In the middle cluster the samples feature both quartz and D/G peaks, with those that have higher intensity D/G peaks in the bottom of the cluster.", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/007-raman-pca.R"), echo = FALSE}
knitr::include_graphics("../figures/raman-pca-with-photos.png", error = FALSE)
```

```{r xrf-si, cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/008-xrf-violin-plot.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/008-xrf-violin-plot.R", encoding = "UTF-8")))

#extract samples above 80% Si 
Si_90 <- Points.xrf %>% 
  filter(`Si` > 90)
#total number of samples in dataset

#calculate percentage of samples above 80% Si
#si_perc <- round(((length(unique(Si_80$sample_id))/length(unique(Points.xrf$sample_id)))*100), 0)
si_90 <- length(unique(Si_90$sample_id))
```

```{r xrf-pca-var, cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/009-xrf-pca-alt-impute.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/009-xrf-pca.R", encoding = "UTF-8")))

@@

```{r xrf-violin-plot, fig.cap= "Violin plot of the XRF values per element. Median and interquartile range represented with a boxplot, surrounded by a kernel density estimation. Values above LOD is noted by *n*. The bimodal distribution of Si is visible in the plot, as well as the low number of values above LOD for V, Y and Zn.", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/008-xrf-violin-plot.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/008-xrf-violin-plot.R", encoding = "UTF-8")))
knitr::include_graphics("../figures/008-xrf-violin-plot.png", error = FALSE)
```

\linebreak
<!--
```{r xrf-pca-scree, fig.cap= "xrf loadings", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/009-xrf-pca-scree.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/009-xrf-pca-scree.R", encoding = "UTF-8")))
knitr::include_graphics("../figures/009-xrf-pca-scree.png", error = FALSE)
```

\linebreak
 
```{r xrf-pca-loadings, fig.cap= "xrf scree plot", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/009-xrf-pca-loadings.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/009-xrf-pca-loadings.R", encoding = "UTF-8")))
knitr::include_graphics("../figures/009-xrf-pca-loadings.png", error = FALSE)
```
-->

@@

```{r xrf-pca, fig.cap= "XRF PCA score plots: a) Score plot featuring the full dataset, cluster 1 includes all samples with Si > 90%, while the remaining samples seem to cluster somewhat based on Fe and Zr content. This results in most of the darker materials clustering towards the bottom of the plot. b) Score plot with > 90% Si samples excluded. Similar to plot A there is a separation based Fe and Zr values, with darker material towards the bottom of the plot. Light-to-colourless samples also remain largely together in cluster 3. Cluster 1 stands out by featuring slightly lower Si, Fe, Ti, and Zr values, but higher values in Al, Ca, K, and Mg.", echo = FALSE}
#quietly(eval(parse("./analysis/paper/009-xrf-pca.R", encoding = "UTF-8")))
knitr::include_graphics("../figures/009-xrf-pca-with-photos.png", error = FALSE)
```

@@

```{r xrf-kmeans-boxplot, fig.cap= "Boxplot of XRF values (%) for model A, grouped by K-means cluster. Note that cluster 1 feature all those samples with a Si value above 90%, with cluster 2 featuring higher Fe/Ti/Zr values.", echo = FALSE}
#quietly(eval(parse("./analysis/paper/009-xrf-kmeans-boxplot.R", encoding = "UTF-8")))
knitr::include_graphics("../figures/009-xrf-kmeans-boxplot.png", error = FALSE)
```

@@

```{r nir-xrf-kmeans, fig.cap= "NIR PCA score plot for PC 1-2, with the K-means cluster from XRF PCA model B visualised in the colour of the points. White points are samples that were excluded in the XRF PCA. The trend separating dark-light samples in the NIR PCA mostly keeps cluster 2-3 separated, showing some overlap in the results, but cluster 1 is fairly evenly distributed in the score plot. This means that whatever differences can be discerned in the compositional data does not seem to be detectable using the ASD.", echo = FALSE}
knitr::include_graphics("../figures/009-nir-pca-xrf-kmeans.png", error = FALSE)
```

@@

```{r nir-sample-comparison, fig.cap= "Comparison of three dark samples from separate K-cluster groups (model B). Sample 391 from cluster 3 stands out the most with visible differences in both the NIR and Raman spectra, when compared to the other samples. The main difference between samples 386 and 404 seems to mainly be in the XRF data.", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/010-sample-comparison.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/010-nir-sample-comparison.R", encoding = "UTF-8")))
knitr::include_graphics("../figures/010-nir-sample-comparison.png", error = FALSE)
```

<!--# O2PLS
 \linebreak
 \linebreak

The cross-validation was performed using the *crossval_o2m_adjR2()* function from the *OmicsPLS* package, and based on the results it was decided to use 2 joint components, 3 X specific (NIRS block), and 4 Y specific (XRF block). The data was then fitted to the O2PLS model, with summary statistics of the joint and orthogonal variation in table \@ref(tab:o2pls-summary).

@@

```{r o2pls-summary, fig.width=8, fig.height=10, cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/010-O2PLS-model-fit.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/010-O2PLS-model-fit.R", encoding = "UTF-8")))
#setup a tempd data.frame with the image of the table, and then us kable to give it a table caption
temp.df = data.frame(image="![](../figures/010-O2PLS-model-fit.png)")
temp.mat <- as.matrix(temp.df)
colnames(temp.mat) <- NULL
knitr::kable(temp.mat, caption="A summary of the O2PLS estimated variation for the X (NIRS) and Y (XRF) data")
```

```{r o2pls-plot, fig.cap= "o2pls score plot", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/010-O2PLS.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/010-O2PLS.R", encoding = "UTF-8")))
knitr::include_graphics("../figures/010-O2PLS.png", error = FALSE)
```
-->

@@

```{r nir-pca-loadings, fig.width=20, fig.height=20, fig.cap= "PCA loadings for the first 3 principal components for the NIR data. ", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/004-nir-pca-loadings.R"), echo = FALSE}
quietly(source("./analysis/paper/004-nir-pca-loadings.R", encoding = "UTF-8"))
knitr::include_graphics("../figures/004-nir-pca-loadings.png", error = FALSE)
```

@@

```{r raman-pca-loadings, fig.cap= "PCA loadings for the first 3 principal components for the Raman data", cache=TRUE, cache.extra=tools::md5sum("./analysis/paper/007-raman-pca-loadings.R"), echo = FALSE}
quietly(eval(parse("./analysis/paper/007-raman-pca-loadings.R", encoding = "UTF-8")))
knitr::include_graphics("../figures/007-raman-pca-loadings.png", error = FALSE)
```

@@

```{r xrf-pca-loadings-scree, fig.cap= "PCA Loadings and eigenvalues for the XRF data. a) eigenvalues and loadings for the full dataset, b) eigenvalues and loadings with samples measuring > 90% ~14~Si excluded.", echo = FALSE}
knitr::include_graphics("../figures/009-xrf-pca-load-scree.png", error = FALSE)
```
\newpage

>>>>

#Import descriptive metadata
metadata.csv <-
  read.csv2(here::here("analysis", "data", "raw_data", "metadata.csv"), sep = ";", header = TRUE, na = c("", "NA", "NULL"), encoding = "UTF-8")

@@

```{r distribution-map, fig.width=15, fig.height=15, fig.fullwidth=TRUE, fig.cap="Geographical distribution of the sampled sites along the river of Ångermanälven, Västerbotten, Sweden.The glacial disturbance is visible in the widely distributd moraine. Made with ©SGU, Lantmäteriet (CC0), Natural Earth, and Riksantikvarieämbetet."}
knitr::include_graphics(here::here("analysis", "figures", "site_distribution_soils_annotated.png"), error = FALSE)
```

@@

```{r sample-hue-pictures, fig.cap= "An assortment of points demonstrating the variation in color and heterogeneity of the material."}
knitr::include_graphics(here::here("analysis", "figures", "sample_hue.jpg"), error = FALSE)
```

@@

\pagebreak

@@

```{r nir-spectra-summary, fig.cap= "Summary of the raw spectra for each colour group. Note the differences in absorbance between the groups, with the dark group featuring a larger range of absorbance values as well as less distinct peaks.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "001-nir-spectra-summary.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "001-nir-spectra-summary.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "001-nir-spectra-summary.png"), error = FALSE)
```

@@

```{r nir-savitzky-transform, fig.cap= "Comparison of NIR spectra between two samples from the dark group. a) the non-processed spectra for sample 258 and 411, b) sample 258 spectra with 2^nd^ derivative Savitzky-Golay filtering, c) sample 411 spectra with 2^nd^ derivative Savitzky-Golay filtering.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "003-nir-savitzky-transform.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "003-nir-savitzky-transform.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "003-nir-savitzky-transform.png"), error = FALSE)
```

@@

```{r nir-savitzky-transform-alt, fig.cap= "SG transformed spectra contrasting the results between the spectra in the dark (black) and light (orange) groups.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "003-nir-savitzky-transform-alt.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "003-nir-savitzky-transform-alt.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "003-nir-savitzky-transform-alt.png"), error = FALSE)
```

```{r nir-peak-summary, fig.cap= "The main features in the NIR spectra can be related to overtones and vibration combinations of OH groups. The 1 400 nm and 1 900 nm peaks are indicative of molecular water, and common in quartz-based materials. The 2 200 nm and 2 350 nm bands indicate metal-OH combinations, where a more intense 2 200 nm band has typically been related to the presence of AlOH, and a more intense 2 350 nm band MgOH. Although difficult to detect in the non-processed spectra, weaker features also appear at 1 065 nm and 1 300 nm, that may be the result of ferrous iron (Fe^2+^)", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "002-nir-peak-summary.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "002-nir-peak-summary.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "002-nir-peak-summary.png"), error = FALSE)
```

@@

```{r nir-pca, fig.width=20, fig.height=20, fig.cap= "PCA score plot", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "004-nir-pca.R")), encoding = "UTF-8", echo = FALSE}
quietly(source(here::here("analysis", "paper","004-nir-pca.R"), encoding = "UTF-8"))
nir.pc1var <- nir.pc1var
nir.pc2var <- nir.pc2var
nir.pc3var <- nir.pc3var
```

@@

```{r nir-pca-fig, fig.width=20, fig.height=20, fig.cap= "NIR PCA score plot for PC 1-3: a) dark and light material is separated out along PC 1, due to absorbance, with white and colourless material separating along PC 2, b) PC 3 adds little new information, mainly separating out a couple of light samples.", echo = FALSE}
#quietly(source("./analysis/paper/004-nir-pca-alt.R", encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "004-nir-pca-with-photos.png"), error = FALSE)
```

@@

```{r raman-spectra-summary, fig.cap= "The non-processed Raman spectra visualised by colour group. Note that the dark material in general feature more intense peaks at wavenumber ~ 1 300 - 1 600 (cm^-1^).", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "005-raman-spectra-summary.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "005-raman-spectra-summary.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "005-raman-spectra-summary.png"), error = FALSE)
```

@@

```{r raman-peak-ident, fig.cap= "The two features that show up in the Raman spectra can be related to quartz and carbon structures in the material. The initial three peaks in the 100 - 500 cm^-1^ range correspond to some of the more well-known vibration modes of quartz. The 1 287 cm-^-1^ and 1 598 cm-^-1^ peaks fall within what is referred to as the D and G band respectively, indicating the presence of carbon structures.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "006-raman-peak-identification.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "006-raman-peak-identification.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "006-raman-peak-identification.png"), error = FALSE)
```

@@

```{r raman-pca-var, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "007-raman-pca.R")), echo = FALSE}
quietly(eval(parse(here::here("analysis", "paper", "007-raman-pca.R"), encoding = "UTF-8")))
raman.pc1var <- raman.pc1var
raman.pc2var <- raman.pc2var
raman.pc3var <- raman.pc3var
```

@@

```{r raman-pca, fig.cap= "PCA score plot for Raman data. Objects that feature only quartz peaks are clustered towards the top left of the plot, whilst the darker material that feature dominant D/G peaks are distributeted towards the right of the plot. In the middle cluster the samples feature both quartz and D/G peaks, with those that have higher intensity D/G peaks in the bottom of the cluster.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "007-raman-pca.R")), echo = FALSE}
knitr::include_graphics(here::here("analysis", "figures", "raman-pca-with-photos.png"), error = FALSE)
```

```{r xrf-si, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "008-xrf-violin-plot.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "008-xrf-violin-plot.R"), encoding = "UTF-8"))

#extract samples above 80% Si 
Si_90 <- Points.xrf %>% 
  filter(`Si` > 90)

#calculate percentage of samples above 80% Si
si_90 <- length(unique(Si_90$sample_id))
```

```{r xrf-pca-var, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "009-xrf-pca-alt-impute.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "009-xrf-pca.R"), encoding = "UTF-8"))

@@

```{r xrf-violin-plot, fig.cap= "Violin plot of the XRF values per element. Median and interquartile range represented with a boxplot, surrounded by a kernel density estimation. Values above LOD is noted by *n*. The bimodal distribution of Si is visible in the plot, as well as the low number of values above LOD for V, Y and Zn.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "008-xrf-violin-plot.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "008-xrf-violin-plot.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "008-xrf-violin-plot.png"), error = FALSE)
```

@@

```{r xrf-pca, fig.cap= "XRF PCA score plots: a) Score plot featuring the full dataset, cluster 1 includes all samples with Si > 90%, while the remaining samples seem to cluster somewhat based on Fe and Zr content. This results in most of the darker materials clustering towards the bottom of the plot. b) Score plot with > 90% Si samples excluded. Similar to plot A there is a separation based Fe and Zr values, with darker material towards the bottom of the plot. Light-to-colourless samples also remain largely together in cluster 3. Cluster 1 stands out by featuring slightly lower Si, Fe, Ti, and Zr values, but higher values in Al, Ca, K, and Mg.", echo = FALSE}
knitr::include_graphics(here::here("analysis", "figures", "009-xrf-pca-with-photos.png"), error = FALSE)
```

@@

```{r xrf-kmeans-boxplot, fig.cap= "Boxplot of XRF values (%) for model A, grouped by K-means cluster. Note that cluster 1 feature all those samples with a Si value above 90%, with cluster 2 featuring higher Fe/Ti/Zr values.", echo = FALSE}
knitr::include_graphics(here::here("analysis", "figures", "009-xrf-kmeans-boxplot.png"), error = FALSE)
```

@@

```{r nir-xrf-kmeans, fig.cap= "NIR PCA score plot for PC 1-2, with the K-means cluster from XRF PCA model B visualised in the colour of the points. White points are samples that were excluded in the XRF PCA. The trend separating dark-light samples in the NIR PCA mostly keeps cluster 2-3 separated, showing some overlap in the results, but cluster 1 is fairly evenly distributed in the score plot. This means that whatever differences can be discerned in the compositional data does not seem to be detectable using the ASD.", echo = FALSE}
knitr::include_graphics(here::here("analysis", "figures", "009-nir-pca-xrf-kmeans.png"), error = FALSE)
```

@@

```{r nir-sample-comparison, fig.cap= "Comparison of three dark samples from separate K-cluster groups (model B). Sample 391 from cluster 3 stands out the most with visible differences in both the NIR and Raman spectra, when compared to the other samples. The main difference between samples 386 and 404 seems to mainly be in the XRF data.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "010-sample-comparison.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "010-nir-sample-comparison.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "010-nir-sample-comparison.png"), error = FALSE)
```

@@

@@

```{r nir-pca-loadings, fig.width=20, fig.height=20, fig.cap= "PCA loadings for the first 3 principal components for the NIR data. ", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "004-nir-pca-loadings.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "004-nir-pca-loadings.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "004-nir-pca-loadings.png"), error = FALSE)
```

@@

```{r raman-pca-loadings, fig.cap= "PCA loadings for the first 3 principal components for the Raman data", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis", "paper", "007-raman-pca-loadings.R")), echo = FALSE}
quietly(source(here::here("analysis", "paper", "007-raman-pca-loadings.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "007-raman-pca-loadings.png"), error = FALSE)
```

@@

```{r xrf-pca-loadings-scree, fig.cap= "PCA Loadings and eigenvalues for the XRF data. a) eigenvalues and loadings for the full dataset, b) eigenvalues and loadings with samples measuring > 90% ~14~Si excluded.", echo = FALSE}
knitr::include_graphics(here::here("analysis", "figures", "009-xrf-pca-load-scree.png"), error = FALSE)
```
\newpage
